### 2020-05-04 Case Studies/03_spam_identification.rmd
spam identification case study
spam identification case study

>>>>

computeMsgLLR <- function(words, freqTable) {

   # discard words not in training data
   words <- words[!is.na(match(words, colnames(freqTable)))]

   # Find which words are present
   present <- colnames(freqTable) %in% words

   sum(freqTable["presentLogOdds", present]) +
      sum(freqTable["absentLogOdds", !present])
}

```{r}
testLLR <- sapply(testMsgWords, computeMsgLLR, trainTable)
```

```{r}
tapply(testLLR, testIsSpam, summary)
```

```{r}
results_df <- data.table( score = testLLR, class = testIsSpam )

ggplot(results_df, aes(score, class, fill = class)) +
   geom_boxplot() +
   coord_flip()   
```

```{r}
typeIErrorRate <- function(tau, llrVals, spam) {
   classify <- llrVals > tau
   sum(classify & !spam) / sum(!spam)
}

typeIErrorRate(0, testLLR, testIsSpam)
typeIErrorRate(-20, testLLR, testIsSpam)

error_rates <- sapply(seq(-30, 30, 1), function(cutoff) c(cutoff = cutoff, rate = typeIErrorRate(cutoff, testLLR, testIsSpam)))

er_df <- data.table(t(error_rates))

ggplot(er_df, aes(cutoff, rate)) +
   geom_line(col = "darkblue") +
   labs(title = "False Positive Error Rates")
```

```{r}
typeIErrorRates <- function(llrVals, isSpam) {
   o <- order(llrVals)
   llrVals <- llrVals[o]
   isSpam <- isSpam[o]

   idx <- which(!isSpam)
   N <- length(idx)
   list(error = (N:1)/N, values = llrVals[idx])
}

### Computational Considerations

```{r}
smallNums <- rep((1/2)^40, 2000000)
largeNum <- 10000

print(sum(smallNums), digits = 20)

print(largeNum + sum(smallNums), digits = 20)

for(i in 1:length(smallNums)) {
   largeNum <- largeNum + smallNums[i]
}

print(largeNum, digits = 20)
```

## Recursive Partitioning and Classification Trees

### Revised E-mail Data Structure

```{r}
header <- sampleSplit[[1]]$header

header[1:12]

header[1] = sub("^From", "Top-From:", header[1])

headerPieces <- read.dcf(textConnection(header), all = T)

headerPieces[, "Delivered-To"]

headerVec <- unlist(headerPieces)
dupKeys <- sapply(headerPieces, function(x) length(unlist(x)))
names(headerVec) <- rep(colnames(headerPieces), dupKeys)

headerVec[ which(names(headerVec) == "Delivered-To") ]

length(headerVec)

length(unique(names(headerVec)))
```

```{r}
processHeader <- function(header) {
   # modify the first line to create a key:value pair
   header[1] <- sub("^From", "Top-From:", header[1])

   headerMat <- read.dcf(textConnection(header), all = T)
   headerVec <- unlist(headerMat)

   dupKeys <- sapply(headerMat, function(x) length(unlist(x)))
   names(headerVec) <- rep(colnames(headerMat), dupKeys)

   return(headerVec)
}

headerList <- lapply(sampleSplit,
                     function(msg) {
                        processHeader(msg$header)
                     })

contentTypes <- sapply(headerList, function(header)
   header["Content-Type"])

names(contentTypes) <- NULL

contentTypes
```

#### Attachments Revisited

```{r}
hasAttach <- grep("^ *multi", tolower(contentTypes))
hasAttach

boundaries <- getBoundary(contentTypes[ hasAttach ])
boundaries

boundary <- boundaries[9]
body <- sampleSplit[[15]]$body

range <- diff(c(bStringLocs[-1], eStringLoc))

body[1:range]

processAttach <- function(body, contentType ) {

   boundary <- getBoundary(contentType)

   n <- length(body)

   if(length(bStringLocs) == 2) {

      bodyContent <- body[(bStringLocs[1] + 2):(bStringLocs[2] - 1)]

      emptyLines <- which(bodyContent == "")
      bodyContent <- bodyContent[ - emptyLines]

      attachContent <- body[(bStringLocs[2] + 1):n]

      aLen <- diff(c(bStringLocs[-1], eStringLoc))
      aType <- body[bStringLocs[-1] + 1]

      if(length(aLen) == length(aType)) {      
         attachments <- data.frame(aLen = aLen, aType = aType)
      } else {
         attachments <- data.frame(aLen = c(), aType = c())
      }

   } else {
      if( length(bStringLocs) == 0 ) {
         bodyContent <- body
      } else {
         bodyContent = body
      }
      attachments <- data.frame(aLen = c(), aType = c())
   }

   return(list(body = bodyContent, attachDF = attachments ))
}

#### More E-Mails

bodyList <- lapply(sampleSplit, function(msg) msg$body)
attList <- mapply(processAttach, bodyList[hasAttach],
                  contentTypes[hasAttach], SIMPLIFY = F)

lens <- sapply(attList, function(processedA)
                           processedA$attachDF$aLen)
```

```{r}
readEmail <- function(dirName) {
   # retrieve the names of files in the directory
   fileNames <- list.files(dirName, full.names = T)

   # drop files that are not email
   notEmail <- grep("cmds$", fileNames)

   if( length(notEmail) > 0 ) fileNames = fileNames[ - notEmail ]

   # read all files in the directory
   lapply(fileNames, readLines, encoding = "latin1")
}

processAllEmail <- function(dirName, isSpam = F) {

   # read all files in the directory
   messages <- readEmail(dirName)

   fileNames <- names(messages)
   n <- length(messages)

   # split header from body
   eSplit <- lapply(messages, splitMessage)
   rm(messages)

   # process header as named character vector
   headerList <- lapply(eSplit, function(msg)
                           processHeader(msg$header))

   # extractd content-type key                        
   contentTypes <- sapply(headerList, function(header)
                                       header["Content-Type"])

   # extract the body
   bodyList <- lapply(eSplit, function(msg) msg$body)
   rm(eSplit)

   # which email have attachements
   hasAttach <- grep("^ *multi", tolower(contentTypes))

   # get summary stats for attachments and the shorter body
   attList <- mapply(processAttach, bodyList[hasAttach],
                     contentTypes[hasAttach], SIMPLIFY = F)

   bodyList[hasAttach] <- lapply(attList, function(attEl)
                                             attEl$body)

   attachInfo <- vector("list", length = n)
   attachInfo[ hasAttach ] <- lapply(attList,
                                     function(attEl) attEl$attachDf)

   # prepare return structure
   emailList <- mapply(function(header, body, attach, isSpam) {
      list(isSpam = isSpam, header = header,
           body = body, attach = attach)
   },
   headerList, bodyList, attachInfo,
   rep(isSpam, n), SIMPLIFY = F)

   names(emailList) <- fileNames

   invisible(emailList)
}

```{r}
emailStruct <- mapply(processAllEmail, directories,
                      isSpam = rep( c(F, T), 3:2))
emailStruct <- unlist(emailStruct, recursive = F)
```

```{r}
sampleStruct <- emailStruct[ 1 ]
```

### Deriving Variables from the email Messages
